### 2020-05-15 Organismal_paper/Organismal_manuscript_and_code.Rmd
knitted modified MS text

---
bibliography: manuscript.bib
csl: pnas.csl
site: "bookdown::bookdown_site"
output:
  bookdown::pdf_document2:
    toc: FALSE
indent: TRUE
monofont: "Times New Roman"
fontsize: 12pt
header-includes:
  - \usepackage[left]{lineno}
  - \linenumbers
  - \usepackage{setspace}
  - \doublespacing
---

@@

SSiphonophores have fascinated zoologists for centuries for their extremely subspecialized colonial organization and integration. Today we hold more knowledge than ever on the morphological diversity of this group due to the extensive work of siphonophore taxonomists in the past few decades (cite Phil, Steve, etc), which has been elegantly synthesized in detailed synopsis [@totton1954synopsis, @mapstone2014global]. In addition, recent advances in phylogenetic analyses of siphonophores [@munro2018improved;@damian2019evolution] have provided a macroevolutionary context to interpret this diversity. With these assets in hand, we can now begin to study siphonophores from an orthogonal perspective, focusing on the diversity and evolutionary history of specific structures. Here we focus on one of such structures: the tentilla. Like many cnidarians, siphonophore tentacles bear side branches (tentilla) with nematocysts. But unlike other cnidarians, most siphonophore tentilla are dynamic structures that react to prey encounters by shooting the nematocyst battery to slap around the prey. This maximizes the surface area of contact between the nematocysts and the prey they fire upon. In addition, siphonophore tentilla present a remarkable diversity of morphologies, sizes, and nematocyst complements (Fig, /@ref(figure2)). Our overarching aim is to organize all this phenotypic diversity in a phylogenetic context, and identify the evolutionary processes that generated it.

@@

Nematocysts are unique biological weapons for defense and prey capture exclusive to the phylum Cnidaria. Mariscal ([@mariscal1974nematocyst]) reported that hydrozoans have the largest diversity of nematocyst types among cnidarians. Among them, siphonophores present the greatest variety of types [@mapstone2014global], and vary widely across taxa in which and how many types they carry on their tentacles. While recent descriptive studies have expanded and confirmed our understanding of this diversity, the evolutionary history of nematocyst type gain and loss in siphonophores remains unexplored. Thus, here we reconstruct the evolution of shifts, gains, and losses of nematocyst types, subtypes, and other major categorical traits that led to the extant diversity we see in siphonophore tentilla. In [@damian2019evolution], we found strong associations between piscivory and haploneme shape. These associations could have been produced by convergent changes in the adaptive optima of these characters. Here we set out to test this hypothesis using comparative model fitting methods. Analyzing the diversity of morphological states from a phylogenetic perspective allows us to identify the specific evolutionary processes that gave rise to it. Here we fit and compare a variety of macroevolutionary models to siphonophore tentilla morphology measurement data to identify instances of neutral divergence, stabilizing selection, changes in the speed of evolution, convergent evolution, and paedomorphosis (larval characters present in mature colonies).

@@

When looking at the morphospace of species in different feeding guilds, we also used PCA on the complete tentacular character dataset transforming inapplicable states of absent characters to zeros to account for similarity based on character presence/absence. Using these principal components, we examined the occupation of the morphospace across species in different feeding guilds using a phylogenetic MANOVA with the package *geiger* [@harmon2008geiger] to assess the variation explained, and a morphological disparity test with the package *geomorph* [@adams2016geomorph] to assess differences in the extent occupied by each guild.

@@

In order to observe the discharge behavior of different tentilla, we recorded high speed footage (1000-3000 fps) of tentillum and nematocyst discharge by live siphonophore specimens (26 species) using a Phantom Miro 320S camera mounted on a stereoscopic microscope. We mechanically elicited tentillum and nematocyst discharge using a fine metallic pin. We used the Phantom PCC software to analyze the footage. For the 10 species recorded, we measured total cnidoband discharge time (ms), heteroneme filament length (μm), and discharge speeds (mm/s) for cnidoband, heteronemes, haplonemes, and heteroneme shafts when possible (data available in the Dryad repository).

>>>>

---
bibliography: org_manuscript.bib
csl: pnas.csl
site: "bookdown::bookdown_site"
output:
  bookdown::pdf_document2:
    toc: FALSE
indent: TRUE
monofont: "Times New Roman"
fontsize: 12pt
header-includes:
  - \usepackage[left]{lineno}
  - \linenumbers
  - \usepackage{setspace}
  - \doublespacing
---

@@

SSiphonophores have fascinated zoologists for centuries for their extremely subspecialized colonial organization and integration. Today we hold more knowledge than ever on the morphological diversity of this group due to the extensive work of siphonophore taxonomists in the past few decades (cite Phil, Steve, etc), which has been elegantly synthesized in detailed synopsis [@totton1965synopsis, @mapstone2014global]. In addition, recent advances in phylogenetic analyses of siphonophores [@munro2018improved;@damian2019evolution] have provided a macroevolutionary context to interpret this diversity. With these assets in hand, we can now begin to study siphonophores from an orthogonal perspective, focusing on the diversity and evolutionary history of specific structures. Here we focus on one of such structures: the tentilla. Like many cnidarians, siphonophore tentacles bear side branches (tentilla) with nematocysts. But unlike other cnidarians, most siphonophore tentilla are dynamic structures that react to prey encounters by shooting the nematocyst battery to slap around the prey. This maximizes the surface area of contact between the nematocysts and the prey they fire upon. In addition, siphonophore tentilla present a remarkable diversity of morphologies, sizes, and nematocyst complements (Fig, /@ref(figure2)). Our overarching aim is to organize all this phenotypic diversity in a phylogenetic context, and identify the evolutionary processes that generated it.

@@

Nematocysts are unique biological weapons for defense and prey capture exclusive to the phylum Cnidaria. Mariscal ([@mariscal1974nematocysts]) reported that hydrozoans have the largest diversity of nematocyst types among cnidarians. Among them, siphonophores present the greatest variety of types [@mapstone2014global], and vary widely across taxa in which and how many types they carry on their tentacles. While recent descriptive studies have expanded and confirmed our understanding of this diversity, the evolutionary history of nematocyst type gain and loss in siphonophores remains unexplored. Thus, here we reconstruct the evolution of shifts, gains, and losses of nematocyst types, subtypes, and other major categorical traits that led to the extant diversity we see in siphonophore tentilla. In [@damian2019evolution], we found strong associations between piscivory and haploneme shape. These associations could have been produced by convergent changes in the adaptive optima of these characters. Here we set out to test this hypothesis using comparative model fitting methods. Analyzing the diversity of morphological states from a phylogenetic perspective allows us to identify the specific evolutionary processes that gave rise to it. Here we fit and compare a variety of macroevolutionary models to siphonophore tentilla morphology measurement data to identify instances of neutral divergence, stabilizing selection, changes in the speed of evolution, convergent evolution, and paedomorphosis (larval characters present in mature colonies).

@@

When looking at the morphospace of species in different feeding guilds, we also used PCA on the complete tentacular character dataset transforming inapplicable states of absent characters to zeros to account for similarity based on character presence/absence. Using these principal components, we examined the occupation of the morphospace across species in different feeding guilds using a phylogenetic MANOVA with the package *geiger* [@harmon2007geiger] to assess the variation explained, and a morphological disparity test with the package *geomorph* [@adams2016geomorph] to assess differences in the extent occupied by each guild.

@@

In order to observe the discharge behavior of different tentilla, we recorded high speed footage (1000-3000 fps) of tentillum and nematocyst discharge by live siphonophore specimens (26 species) using a Phantom Miro 320S camera mounted on a stereoscopic microscope. We mechanically elicited tentillum and nematocyst discharge using a fine metallic pin. We used the Phantom PCC software to analyze the footage. For the 10 species recorded, we measured total cnidoband discharge time (ms), heteroneme filament length ($\mu$m), and discharge speeds (mm/s) for cnidoband, heteronemes, haplonemes, and heteroneme shafts when possible (data available in the Dryad repository).
